---
title: "HW2"
author: Yi Sun
output: github_document
---

Solution to HW2.

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Define a path to the Mr. Trash Wheel dataset.

```{r}
path_trash_wheel = "./Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```


Read and clean the Mr. Trash Wheel dataset

```{r}
trashwheel_df = 
  read_excel(
    path = path_trash_wheel, 
    sheet = "Mr. Trash Wheel", 
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read and clean the 2017 and 2018 precipitation datasets.

```{r}
precipitation_2017 = 
  read_excel(
    path = path_trash_wheel, 
    sheet = "2017 Precipitation", 
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)

precipitation_2018 = 
  read_excel(
    path = path_trash_wheel, 
    sheet = "2018 Precipitation", 
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```

Next, combine the 2017 and 2018 precipitation datasets with the helper dataframe `month_df`.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precipitation_combine = 
  bind_rows(precipitation_2017, precipitation_2018)

precipitation_combine = 
  left_join(precipitation_combine, month_df, by = "month")
```

There are `r nrow(trashwheel_df)` rows in the final Mr. Trash Wheel dataset and `r nrow(precipitation_combine)` rows in the combined precipitation dataset of 2017 and 2018. 

The final Mr. Trash Wheel dataset contains information on year, month, date, weight and volume of trash collected, and quantity of some specific kinds of trash. 

The combined precipitation dataset of 2017 and 2018 contains precipitation data by month for 2017 and 2018. 

* The total precipitation in 2018 was `r precipitation_combine %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.
* The median number of sports balls in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.

## Problem 2

Define a path to the NYC subway transit dataset.

```{r}
path_NYC_subway = "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv"
```

Read and clean the NYC subway transit dataset, only retaining variables needed.

```{r}
NYC_subway_df = 
  read_csv(path_NYC_subway) %>% 
  janitor::clean_names() %>% 
  select(line:entrance_type, entry, vending, ada) %>% 
  mutate(entry = as.logical(
    ifelse(entry == "YES", 1, 0)
  ))
```

The resulting dataset of NYC subway transit `NYC_subway_df` only keeps variables like line, station name, latitude and longitude of station, routes served, entry, vending, entrance type, and ADA compliance.

The variable `entry` is converted to a logical variable.

There are `r nrow(NYC_subway_df)` rows and `r ncol(NYC_subway_df)` columns in the resulting dataset of NYC subway transit `NYC_subway_df`.

The data in `NYC_subway_df` are still not tidy. There are still redundant columns for routes served, or the route numbers that should be values are still in the variable names.

* There are `r NYC_subway_df %>% distinct(line, station_name) %>% nrow()` distinct stations.
* There are `r NYC_subway_df %>% filter(ada == TRUE) %>% distinct(line, station_name) %>% nrow()` ADA compliant stations.
* `r nrow(filter(NYC_subway_df, entry == TRUE & vending == "NO"))/nrow(filter(NYC_subway_df, vending == "NO"))*100`% of station entrances / exits without vending allow entrance.

Reformat `NYC_subway_df` to make route number and route name distinct variables.

```{r}
NYC_subway_df = 
  NYC_subway_df %>% 
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
```

* `r NYC_subway_df %>% filter(route_name == "A") %>% distinct(line, station_name) %>% nrow()` distinct stations serve the A train.
* Of the stations that serve the A train, `r NYC_subway_df %>% filter(route_name == "A" & ada == TRUE) %>% distinct(line, station_name) %>% nrow()` distinct stations are ADA compliant.

## Problem 3

Define a path to the pols-month dataset.

```{r}
path_to_pols_month = "./data_538/pols-month.csv"
```

Read and clean the pols-month dataset.

```{r}
pols_month_df = 
  read_csv(path_to_pols_month) %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day"), "-", convert = TRUE) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(president = ifelse(prez_gop == 1, "gop", "dem")) %>% 
  select(-c(prez_gop, prez_dem))
```

