---
title: "HW2"
author: Yi Sun
output: github_document
---

Solution to HW2.

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Define the path to the Mr. Trash Wheel dataset.

```{r}
path_trash_wheel = "./Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```


Read and clean the Mr. Trash Wheel dataset

```{r}
trashwheel_df = 
  read_excel(
    path = path_trash_wheel, 
    sheet = "Mr. Trash Wheel", 
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read and clean the 2017 and 2018 precipitation datasets.

```{r}
precipitation_2017 = 
  read_excel(
    path = path_trash_wheel, 
    sheet = "2017 Precipitation", 
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)

precipitation_2018 = 
  read_excel(
    path = path_trash_wheel, 
    sheet = "2018 Precipitation", 
    skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)
```

Next, combine the 2017 and 2018 precipitation datasets with the helper dataframe `month_df`.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precipitation_combine = 
  bind_rows(precipitation_2017, precipitation_2018)

precipitation_combine = 
  left_join(precipitation_combine, month_df, by = "month")
```

There are `r nrow(trashwheel_df)` rows in the final Mr. Trash Wheel dataset and `r nrow(precipitation_combine)` rows in the combined precipitation dataset of 2017 and 2018. 

The final Mr. Trash Wheel dataset contains information on year, month, date, weight and volume of trash collected, and quantity of some specific kinds of trash. 

The combined precipitation dataset of 2017 and 2018 contains precipitation data by month for 2017 and 2018. 

* The total precipitation in 2018 was `r precipitation_combine %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.
* The median number of sports balls in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.